Multi-target detection and clustering with application to cryo-electron microscopy

### 2019-02-11 inverse_problems/inverse_problems.tex
no message

If $M_i$ denotes the number of hits at location $i$,  then by definition of the Poisson process $M_i$'s are i.i.d.\ and $M_i \sim \Poisson(\gamma)$. Conditional on the value of $M = (M_1,\dots,M_{N-L+1})$, if we let $X_1^{i},\dots,X_{M_i}^i$ denote the random vectors with position 0 located at $i$, then $X_{k_1}^{i}$ and $X_{k_2}^{i}$ are independent for $k_1 \ne k_2$.
With this notation, we can write each entry as:
%
\begin{align}
%
Y[i] = \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} X_k^{i-j}[j].
%
\end{align}

@@

\subsection{Equivalence between the spacing condition and the Poisson process} \label{sec:poisson}

In this section we omit the affect of the noise as it remains the same for the well-separated model.

Let us denote by $m_Y^q$ the moments of $Y$:
%
\begin{align}
%
m_y^1[i] = \E X[i], \quad 0 \le i \le L-1,
%
\end{align}
%
\begin{align}
%
m_y^2[i,j] = \E X[i] X[j], \quad 0 \le i,j \le L-1,
%
\end{align}
%
and
%
\begin{align}
%
m_y^3[i,j,k] = \E X[i] X[j] X[k], \quad 0 \le i,j,k \le L-1.
%
\end{align}

%We will also denote by $\L_l$ the autocorrelations of $X$, $l = 1,2,3$:
%%
%\begin{align}
%%
%\L_1 = \sum_{i=0}^{L-1} \M_1[i] \in \R,
%%
%\end{align}
%%
%\begin{align}
%%
%\L_2(\Delta) = \sum_{i=0}^{L-1} \M_2[i,i+\Delta]  \in \R^L,
%%
%\end{align}
%%
%and
%%
%\begin{align}
%%
%\L_3(\Delta_1,\Delta_2) = \sum_{i=0}^{L-1} \M_3[i,i+\Delta_1,i+\Delta_2]  \in \R^{L \times L}.
%%
%\end{align}
%
%
%Note that in the strongly-separated model, the first three observed moments are, respectively, $\L_1$, $\L_2(\Delta)$, and $\L_3(\Delta_1,\Delta_2)$.
%
%%

@@

\begin{proposition} \label{prop:poisson}
    For each $i$,  the moments of the Poisson process are equal to 
    \begin{align}
    m_y^1[i] &= \gamma a_X^1 \nonumber \\
    m_y^2[i,i+\ell] &= (\gamma a_X^1)^2 + \gamma a_X^2[\ell],\quad \ell=0,\ldots,L-1, \\
    m_y^3[i,i+\ell_1,i+\ell_2] &= (\gamma a_X^1)^3 + \gamma a_X^1  \cdot ( \gamma a_X^2[\ell_1] + \gamma a_X^2[\ell_2] + \gamma a_X^2[\ell_2-\ell_1]) + \gamma a_X^3[\ell_1,\ell_2]. \nonumber  
    \end{align}
\end{proposition}

@@

Autocorrelations in $d$ dimensions is defined for  $\ell_1,\ldots,\ell_{q-1}\in\mathbb{Z}^d$  as 
\begin{equation}
a_Z^q[\ell_1,\ldots,\ell_{q-1}]   = \E\left\{\frac{1}{m^d} \sum_{i\in\mathbb{Z}^d} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}]\right\}.
\label{eq:ac_d_dimension}
\end{equation}
Interestingly, for dimensions greater than one almost all deterministic (that is, the distribution of $Z$ is a Delta function) signals are determined uniquely from their second-order autocorrelation, up to two symmetries: sign (or phase for complex signals) and reflection through the origin (with conjugation in the complex case)~\cite{hayes1982reconstruction}. 
This is not true for 1-D signals: generally there are $2^{L-2}$ signals with the same second-order autocorrelation (after eliminating the symmetries)~\cite{beinert2015ambiguities,bendory2017fourier}.

In the literature, the phase and reflection symmetries are usually called trivial ambiguities. If the mean of signal is available and non-zero, the sign symmetry can be resolved. However, determining the reflection symmetry still requires additional information, beyond the second-order autocorrelation.

@@

Clearly, the first-order autocorrelation provides us one equation. For second-order autocorrelations, notice  that $a_y^2[\ell]$ suffers no noise-induced bias for $\ell$ in $1$ to $L-1$. Thus if $\sigma$ is known we obtain $L$ equations, while if not we omit $\ell = 0$. This has the practical effect that we need not know $\sigma$ to make sense of the computed quantities, while loosing one one equation. 
Likewise, for third-order autocorrelations, $a_y^3[\ell_1, \ell_2]$ for $0 \leq \ell_1, \ell_2 \leq L-1$ such that $\ell_2 \leq \ell_1$ includes all relevant entries for our purpose (this accounts for symmetries): $\frac{(L+1)(L+2)}{2}-2$ in total. 
If we further exclude any such that $\ell_1, \ell_2$ or $\ell_1 - \ell_2$ are zero to avoid the need to estimate $\sigma$---there are $\frac{(L-1)(L-2)}{2}$ remaining entries. 
Hence, if $\sigma$ is know we have 
\begin{align*}
1 + L + \frac{(L+1)(L+2)}{2}-2 = \frac{1}{2} L (L+5)
\end{align*}
and if it unknown
\begin{align*}
1 + (L-1) + \frac{(L-1)(L-2)}{2} = \frac{1}{2} L (L-1) + 1
\end{align*}
coefficients in total. Since we aim to estimate $KL$ parameters (for the $K$ signals of length $L$) plus $K$ parameters (for the densities $\gamma_k$), an absolute upper bound on $K$  is
\begin{equation}
K\leq \frac{L(L+5)}{2(L+1)},
\end{equation}
\begin{align*}
K \leq \frac{ L (L-1) + 1}{2(L+1)}.
\end{align*}
Thus, we conjecture that approximately $L/2$ signals can be decomposed from  the first three mixed autocorrelations.

@@

\subsection{The first moment of $Y$}

To compute the first moment of $Y$, we will first condition on $M = (M_1,\dots,M_{N-L+1})$, and then average over $M$. We have:
%
\begin{align}
%
\E[Y[i] | M] = \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} \E X_k^{i-j}[j]
= \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} m_y^1[j]
= \sum_{j=0}^{L-1} M_{i-j} m_y^1[j].
%
\end{align}
%
Now taking expectations over $M$ we see:
%
\begin{align}
%
\E Y[i] = \gamma \sum_{j=0}^{L-1}  m_y^1[j] = \gamma a_x^1.
%
\end{align}

@@

\subsection{The second moment of $Y$}

@@

\subsection{The third moment of $Y$}

>>>>

If $M_i$ denotes the number of hits at location $i$,  then by definition of the Poisson process $M_i$'s are i.i.d.\ and $M_i \sim \Poisson(\gamma)$. Conditional on the value of $M = (M_1,\dots,M_{N-L+1})$, if we let $X_1^{i},\dots,X_{M_i}^i$ denote the random vectors with position 0 located at $i$, then $X_{k_1}^{i}$ and $X_{k_2}^{i}$ are independent for $k_1 \ne k_2$.
With this notation, we can write each entry as:
%
\begin{align}
%
y[i] = \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} X_k^{i-j}[j].
%
\end{align}

@@

\subsection{Equivalence between the autocorrelations of under the separation and the Poisson process} \label{sec:poisson}

In this section we omit the affect of the noise as it remains the same for the well-separated model. We aim to show that the moments of

Let us denote by $m_Y^q$ the moments of $Y$:
%
\begin{align}
%
m_X^1[i] &= \E X[i], \quad 0 \le i \le L-1,   \nonumber\\ 
m_X^2[i,j] &= \E X[i] X[j], \quad 0 \le i,j \le L-1, \\ 
m_X^3[i,j,k] &= \E X[i] X[j] X[k], \quad 0 \le i,j,k \le L-1. \nonumber
\end{align}
%
Note that 
\begin{align}
a_X^1 &= \frac{1}{L}\sum_{i=0}^{L-1}m_X^1[i], \nonumber \\
a_X^2[\ell] &= \frac{1}{L}\sum_{i=0}^{L-1}m_X^2[i,i+\ell], \\
a_X^3[\ell_1,\ell_2] &= \frac{1}{L}\sum_{i=0}^{L-1}m_X^3[i,i+\ell_1,i+\ell_2]. \nonumber
\end{align}

@@

\begin{proposition} \label{prop:poisson}
    For any $i$,  the moments under the Poisson process are equal to 
    \begin{align}
    m_y^1[i] &= \gamma L a_X^1 \nonumber \\
    m_y^2[i,i+\ell] &= (\gamma a_X^1)^2 + \gamma a_X^2[\ell],\quad \ell=0,\ldots,L-1, \\
    m_y^3[i,i+\ell_1,i+\ell_2] &= (\gamma a_X^1)^3 + \gamma a_X^1  \cdot ( \gamma a_X^2[\ell_1] + \gamma a_X^2[\ell_2] + \gamma a_X^2[\ell_2-\ell_1]) + \gamma a_X^3[\ell_1,\ell_2]. \nonumber  
    \end{align}
\end{proposition}

@@

Autocorrelations in $d$ dimensions is defined for  $\ell_1,\ldots,\ell_{q-1}\in\mathbb{Z}^d$  as 
\begin{equation}
a_Z^q[\ell_1,\ldots,\ell_{q-1}]   = \E\left\{\frac{1}{m^d} \sum_{i\in\mathbb{Z}^d} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}]\right\}.
\label{eq:ac_d_dimension}
\end{equation}
Interestingly, for dimensions greater than one almost all deterministic (that is, the distribution of $Z$ is a Delta function) signals are determined uniquely from their second-order autocorrelation, up to two symmetries: sign (or phase for complex signals) and reflection through the origin (with conjugation in the complex case)~\cite{hayes1982reconstruction}. 
If the mean of signal is available and non-zero, the sign symmetry can be resolved. However, determining the reflection symmetry still requires additional information, beyond the second-order autocorrelation.
The case of 1-D signals is essentially different: generally there are $2^{L-2}$ signals with the same second-order autocorrelation (after eliminating  symmetries)~\cite{beinert2015ambiguities,bendory2017fourier}.

@@

\TODO{Probably should be re-written}
Clearly, the first-order autocorrelation provides us one equation. For second-order autocorrelations, notice  that $a_y^2[\ell]$ suffers no noise-induced bias for $\ell$ in $1$ to $L-1$. Thus if $\sigma$ is known we obtain $L$ equations, while if not we omit $\ell = 0$. This has the practical effect that we need not know $\sigma$ to make sense of the computed quantities, while loosing one one equation. 
Likewise, for third-order autocorrelations, $a_y^3[\ell_1, \ell_2]$ for $0 \leq \ell_1, \ell_2 \leq L-1$ such that $\ell_2 \leq \ell_1$ includes all relevant entries for our purpose (this accounts for symmetries): $\frac{(L+1)(L+2)}{2}-2$ in total. 
If we further exclude any such that $\ell_1, \ell_2$ or $\ell_1 - \ell_2$ are zero to avoid the need to estimate $\sigma$---there are $\frac{(L-1)(L-2)}{2}$ remaining entries. 
Hence, if $\sigma$ is know we have 
\begin{align*}
1 + L + \frac{(L+1)(L+2)}{2}-2 = \frac{1}{2} L (L+5)
\end{align*}
and if it unknown
\begin{align*}
1 + (L-1) + \frac{(L-1)(L-2)}{2} = \frac{1}{2} L (L-1) + 1
\end{align*}
coefficients in total. Since we aim to estimate $KL$ parameters (for the $K$ signals of length $L$) plus $K$ parameters (for the densities $\gamma_k$), an absolute upper bound on $K$  is
\begin{equation}
K\leq \frac{L(L+5)}{2(L+1)},
\end{equation}
\begin{align*}
K \leq \frac{ L (L-1) + 1}{2(L+1)}.
\end{align*}
Thus, we conjecture that approximately $L/2$ signals can be decomposed from  the first three mixed autocorrelations.

@@

\subsection{First moment}

To compute the first moment of $y$, we will first condition on $M = (M_1,\dots,M_{N-L+1})$, and then average over $M$. We have:
%
\begin{align}
%
\E[y[i] | M] = \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} \E X_k^{i-j}[j]
= \sum_{j=0}^{L-1} \sum_{k=1}^{M_{i-j}} m_y^1[j]
= \sum_{j=0}^{L-1} M_{i-j} m_y^1[j].
%
\end{align}
%
Now taking expectations over $M$ we see:
%
\begin{align}
%
\E Y[i] = \gamma \sum_{j=0}^{L-1}  m_y^1[j] = \gamma La_x^1.
%
\end{align}

@@

\subsection{Second moment}

@@

\subsection{Third moment}
